---
title: "Pa4 Presentation"
subtitle: "⚔<br/>with xaringan"
author: "Jukun Zhang"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["rutgers", "rutgers-fonts"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

#Hello

##Load library


```{r,message=FALSE, warning=FALSE}
#| label: Loading library 
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ds4ling)
```

---

##Load the dataset

```{r}
#| label: Loading dataset
#| 
mydata1 <- read.csv("C:/Users/zhang/OneDrive/Desktop/pa4/data_raw/fricatives/fricatives.csv")
print(mydata1)
```

---

##Tidy the dataset

```{r}
#| label: Tidy dataset

mydata1_tidy <- mydata1 |>
  pivot_longer(
    cols = c(s_cog, s_skewness, sh_cog, sh_skewness),  
    names_to = c("sound", "measurement"),              
    names_sep = "_",                                   
    values_to = "value"                                
  )

```

---

##Save the tidy version

```{r}
#| label: Save the tidy version

write.csv(mydata1_tidy, "C:/Users/zhang/OneDrive/Desktop/pa4/data_tidy/mydata1_tidy.csv", row.names = FALSE)
```
---

## Summary table of descriptive statistics

```{r}
#| label: Summary tabel of descriptive statistics

summary_table <- mydata1_tidy |>
  group_by(sound, measurement) |>
  summarise(
    mean = mean(value),
    median = median(value),
    sd = sd(value),
    min = min(value),
    max = max(value),
    .groups = "drop"
  )
knitr::kable(summary_table, digits = 2, caption = "Descriptive Statistics for Fricatives")
```

---

##Make a boxplot of center of gravity as a function of phoneme

```{r, echo=FALSE, fig.width= 10, fig.height=6}
#| label: Make a boxplot of center of gravity as a function of phoneme

mydata1_tidy |>
  filter(measurement == "cog") |>
  ggplot(aes(x = sound, y = value, fill = sound)) +
  geom_boxplot() +
  labs(
    title = "Center of Gravity by Phoneme",
    x = "Phoneme",
    y = "Center of Gravity (Hz)"
  ) +
  theme_minimal()
```

---

##Plot skewness as a function of phoneme. Use a statistical transformation

```{r, echo=FALSE, fig.width= 10, fig.height=6}
#| label: Make a plot of skewness as a function of phoneme

mydata1_tidy |>
  filter(measurement == "skewness") |>
  ggplot(aes(x = sound, y = value, color = sound)) +
  stat_summary(fun = mean, geom = "point", size = 4) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2) +
  labs(
    title = "Mean Skewness by Phoneme",
    x = "Phoneme",
    y = "Skewness"
  ) +
  theme_minimal()
```
---

## Filter 'sound == s' from mydata1_tidy

```{r}
s_data <- mydata1_tidy |>
  filter(sound == "s") |>
  pivot_wider(
    names_from = measurement,
    values_from = value
  )

head(s_data)
```
---

##Fit model and give the summary table
```{r}
model1 <- lm(cog ~ skewness, data = s_data)

summary(model1)
```
---

##Make Scatter plot in (8)

```{r, echo=FALSE, fig.width= 8, fig.height=6}
s_data |>
  ggplot(aes(x = skewness, y = cog)) +
  geom_point(color = "steelblue", size = 2) +
  geom_smooth(method = "lm", se = TRUE, color = "darkred") +
  labs(
    title = "Relationship Between Skewness and Center of Gravity for [s]",
    x = "Skewness",
    y = "Center of Gravity (Hz)"
  ) +
  theme_minimal()
```

---

##Check the model1 diagnostics

```{r, message= FALSE, warning=FALSE, fig.width=10, fig.height=6}
diagnosis(model1)
```

---
##Result of the model1

- We investigated whether skewness predicts center of gravity (CoG) for [s] fricative segments using a linear regression model. From the model summary the model revealed a strong, statistically significant positive relationship between skewness and CoG, β = 1675.6, t(48) = 12.02, p < .001. This indicates that as spectral skewness increases, the center of gravity also increases. The model explained a substantial proportion of the variance, R² = 0.75, suggesting that skewness is a strong predictor of frequency energy distribution in [s].

- Model diagnostics (Figure 2) showed no major violations of assumptions: residuals were approximately normally distributed, homoscedastic, and without extreme leverage points. The residual standard error was 419.4 Hz.

- Model diagnostics indicated that assumptions of linearity, homoscedasticity, and normality were met. Residuals were randomly distributed around zero, approximately normally distributed, and followed the expected Q-Q pattern closely (r = 0.994), suggesting the linear model is appropriate for the data.

---

##Load the assumptions.csv dataset and make a scatterplot

```{r, echo=FALSE, fig.width=7, fig.height=4}
mydata2 <- read.csv("C:/Users/zhang/OneDrive/Desktop/pa4/data_raw/fricatives/assumptions.csv")

mydata2 |>
  ggplot(aes(x = x, y = y)) +
  geom_point(size = 2) +
  labs(
    title = "Scatterplot of Assumptions Dataset",
    x = "X",
    y = "Y"
  ) +
theme_minimal()
```
- Why it would not be appropriate to fit a model to this data: 
The scatterplot of the assumptions.csv data shows a highly nonlinear, complex pattern that appears to loop or spiral, rather than follow any kind of linear trend.  Fitting a linear model here would lead to poor fit, biased estimates, and invalid inferences. 







